Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics

### 2019-08-15 review/response.Rmd

The submission has been evaluated by two reviewers. The manuscript is well written and organized. I have some more philosophical issues in addition to those raised by the two reviewers.

The authors need to convince readers (users) this kind of environment is good for daily EDA for the real world of big data/data science.

**We have added a paragraph to the introduction to explain the approach for higher dimensional problems, as follows:**

**PP can be combined with other dimension reduction methods when p is very high. For example, it can be beneficial to first do principal component analysis prior to PP, especially to remove linear dependencies before searching for other types of association. This is the approach used in Cook, Laa, and Valencia (2018), which explores a 56-dimensional parameter space, by first reducing the number of dimensions to the first six principal components, before applying projection pursuit. PCA was appropriate for this problem because reducing to principal component space removed the linear dependencies while preserving the nonlinear relationships that were interesting to discover. Some projection pursuit indexes do incorporate penalty terms to automate removing noise dimensions. It can also be important to have an efficient PP optimizer, particularly when working with high dimensions, because the search space increases exponentially with dimension.**

@@

5) Why use the 5th/95th percentiles in Table 2 instead of the average/sd?

**The main reason is that it provides a range of typical values of the index, which is useful for scaling the index dynamically in the guided tour. While it is possible to use a mean, and maybe +/- 3sd's, its possible that extreme values could occur making it less useful to scale the index for comparison  to other indexes.**

6) There was some confusion regarding the number of PPIs used in the content. There are eight indexes in Table 2; nine indexes in Fig 7; six indexes in Table 3, and seven indices in Section 4.2.2.

**We are considering seven new index functions. In Section 3, we also include the holes index for comparison, because it is an established index with generally good, focused performance. This is now explained in the end of Section 2.2. The holes index is not included in Table 3 because the focus is on summarising our findings on the new index functions. Because the MINE indexes (MIC, TIC) have very similar behaviour in all assessments, we had combined them for this summary. We have now added another line for TIC, because it doesn't use much extra space, and saves confusion about the differing number of indexes. Finally Section 4.2.2 applies a selection of the new index functions to physics problems.**

@@

 **We have added more explanation in Section 3.1.1 outlining the motivation for why these three structures are interesting for the simulation study.**

Second, the simulation study uses six-dimensional data of which four dimensions contribute nuisance and the other two the intended structure. Implicitly, this is clear, but the description does not make this explicit and uses the general notion of p dimensions instead. Please make this more explicit and also provide some justification that this set-up is adequate for targeting the problem at hand.

**We have added an explicit statement that we fix p=6. As mentioned in the statement p=6 results in two independent planes showing the nuisance distribution and one plane showing the structured distribution. This set-up is adequate for most of our study where we follow interpolated paths either between nuisance only projections or from nuisance views to structured projections.**

Thirdly, the simulations refer to sequences of projections (the horizontal axis in plots in Figures 2, 3, and 4. As I understand these sequences are a consequence of optimizing the corresponding projection pursuit index. How can you then actually control the two different scenarios referred to in section 3.2 and 3.3? Or, do you extract a specified sequence of projections? Moreover, how do you make sure that you have the same sequence length for each trial? Please clarify!

@@

**Captions and the text in Sections 3.2-3 have been changed to make it clearer that Figures 2 and 3 present the trace for a specified sequence of projections (since in the simulation study the nuisance and structured views are known), and Figure 4 is different as in this case the path is indeed the result of optimising an index.**

The optimization methods are further discussed in section 3.4 and depend on the structure present in the data. What is the value of these investigations for more realistic settings in which the structure is not known.

### 2019-08-15 Appendix/appendix.Rmd
Preparing resubmission.

```{r timer, fig.cap="Computational perfoamce for PPIs, using sample sizes 100-10000. Colour indicates the PPI. Because the scagnostics calculation is bundled together, the values are the same for all these indexes, and they are really fast to compute. MINE includes the MIC and TIC indexes, and MINE E are computationally more efficient algorithms for these. These, along with dcor2D, are slower with larger sample sizes.", out.width=".6\\textwidth"}
if(!file.exists("../cache/timer.rda")){
  set.seed(2018)
  grandTour100 <- save_history(sphereData(6,100), grand_tour(2), max=4) %>%
    interpolate() %>%
    as.list()

@@

Figures \ref{fig:neutronStarSPLOM} and \ref{fig:bbhSimulation} show the scatter plot matrices of the gravitational wave datasets considered, see Section \ref{sec:phys} for details.

>>>>

```{r timer, fig.cap="Computational performance for PPIs, using sample sizes 100-10000. Colour indicates the PPI. Because the scagnostics calculation is bundled together, the values are the same for all these indexes, and they are really fast to compute. MINE includes the MIC and TIC indexes, and MINE E are computationally more efficient algorithms for these. These, along with dcor2D, are slower with larger sample sizes.", out.width=".6\\textwidth"}
if(!file.exists("../cache/timer.rda")){
  set.seed(2018)
  grandTour100 <- save_history(sphereData(6,100), grand_tour(2), max=4) %>%
    interpolate() %>%
    as.list()

@@

Figures \ref{fig:neutronStarSPLOM} and \ref{fig:bbhSimulation} show the scatter plot matrices of the gravitational wave datasets considered.

### 2019-08-15 scagnostics.Rmd
Preparing resubmission.

abstract: |
  Projection pursuit is used to find interesting low-dimensional projections of high-dimensional data by optimizing an index over all possible projections. Most indexes have been developed to detect departure from known distributions, such as normality, or to find separations between known groups. Here, we are interested in finding projections revealing potentially complex bivariate patterns, using new indexes constructed from scagnostics and a maximum information coefficient, with a purpose to detect unusual relationships between model parameters describing physics phenomena. The performance of these indexes is examined with respect to ideal behaviour, using simulated data, and then applied to problems from gravitational wave astronomy. The implementation builds upon the projection pursuit tools available in the R package, tourr, with indexes constructed from code in the R packages, scagnostics, minerva and mbgraphic.

@@

<!-- 
Main point of paper:

- Some metrics for exploring high-dimensional data by finding interesting pairs of variables, can be used to find interesting projections, but some have problems. 
- What are the ways to evaluate.
- What adjustments can be done.
- How it can be applied

Organisation of this section:

- Motivation first?
- PP
- Scagnostics
- section needs shortening
-->

@@

<!-- check Perisic and Posse (2012) and recent paper by Austrian guys -->

The driving application is from physics, to aid the interpretation of model fits on experimental results. A physical model can be considered to be a set of $p$ free parameters, that cannot be measured directly and are determined by fitting a set of $q~ (p<q)$ experimental observations, for which predictions can be made once the $p$ parameters are estimated. (Note here, that while we may have analytic expressions for the predictions, this is not always the case and we often have to rely on numerical computation.) Different sets of model parameters ($n$) found to be  compatible with the experimental results within a selected level of confidence yield the data to be examined using projection pursuit. A single prediction can be a complicated function of all of the free parameters, and typically $q \in [100,1000]$ and $p \sim 10$. Current practice is to examine pairs of parameters, or combinations produced by intuition or prior knowledge. This begs the question, whether important nonlinear associations are missed because because they are hidden in linear combinations of more than two variables.

@@

To find appropriate projections pursuit indexes for detecting nonlinear dependencies, the literature on variable selection was a starting point. With high-dimensional data, even plotting all pairs of variables can lead to too many plots, which is what "scagnostics" (@scag, @WW08) were developed to address by providing metrics from which to select the most interesting variable pairs. There are eight scagnostics, of which three ("convex", "skinny" and "stringy") are used here. The question is whether these can be adapted into projection pursuit indexes, to search for unusual features in two-dimensional projections of high-dimensional data. Recent PhD research by @Grimm2016 explored the behavior of scagnostics for selecting variables, and proposed two more that have nicer properties, based on smoothing splines and distance correlation. In addition, two more indexes for measuring dependence have been proposed in the machine learning literature, based on information criteria, maximal and total information coefficient (MIC and TIC) [@Reshef1518], with computationally more efficient versions (MIC_e, TIC_e) [@JMLRv1715308]. These are related to original 1D projection pursuit indexes based on entropy (e.g. @huber85, @JS87). This provides seven current indexes for measuring dependence between two variables, and each is available in an R [@rref] package: scagnostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].

@@

This paper is structured as follows. Section \ref{sec:construct} discusses index construction, and how they can be used in the guided tour. Section \ref{sec:investigate} investigates the behavior of the indexes, explored primarily using tour methods. The new guided tour with these indexes is applied to two examples from gravitational wave astronomy (Section \ref{sec:phys}). The latter two parts are connected in that the application of the new indexes to these problems is the main motivation for the paper, and the simulation study, in the first part, was conducted to better understand the behavior of the indexes in general. The techniques in Section \ref{sec:investigate} define procedures that will be generally useful for researchers developing new projection pursuit indexes to visually assess their behavior. Visual methods to diagnose the index behavior is important because PP is primarily used for visualistion.  The paper finishes with a discussion about the limitations of this work, and the potential future directions.

@@

A comparison between these indexes for the **purpose of variable selection**, but not projection pursuit, was discussed in @Grimm2016. It is only available in German, we we summarise the main findings here. The scagnostics measures are flexible, and calculating the full set of measures provides useful guidance in variable selection. However, they are found to be highly sensitive to outlying points and sample size (as a consequence of the binning). Both splines2D and dcor2D are found to be robust in this respect, but splines2D is limited to functional dependence, while dcor2D is found to take large values only in scenarios with large linear correlation. The mutual information based index functions (MIC< TIC) are found to be flexible, but are sensitive to the sample size and often take relatively large values even when no association is present. A brief comparison of MIC and dcor2D was also provided in @2014arXiv1401.7645S.

In addition to the seven indexes described above, we will also include the holes index available in the `tourr` package, see [@CBC93,@Cook:2007:IDG:1557227]. This serves as a benchmark, demonstrating some desired behavior. The index takes maximum values for a central hole in the distribution of the projected data.

@@

if(!file.exists("cache/indexTable.rda")){
  set.seed(505)
  res <- tableTemp()
  for(ds in c("Pipe", "Sine", "Spiral")){
    for (size in c(100, 1000)){
      for (type in c("structure", "noise")){
        res <- bind_rows(res,getIdx(ds, size, type))
      }
    }
  }
  res <- res %>% arrange(data, size, index, type) %>% select(data, size, index, type, value5, value95)
  save(res, file = "cache/indexTable.rda")
} else {
  load("cache/indexTable.rda")
}
#rescale holes index first
res <- res %>%
  rowwise() %>%
  mutate(value5 = if_else(index=="holes", rescaleHoles(value5), value5)) %>%
  rowwise() %>%
  mutate(value95 = if_else(index=="holes", rescaleHoles(value95), value95))
resp <- res %>% filter(size == 1000, data == "Pipe") %>%
  select(index, type, value5, value95) %>%
  rename(lower=value5, upper=value95)
resi <- res %>% filter(size == 1000, data == "Sine") %>%
  select(index, type, value5, value95) %>%
  rename(lower=value5, upper=value95)
ress <- res %>% filter(size == 1000, data == "Spiral") %>%
  select(index, type, value5, value95) %>%
  rename(lower=value5, upper=value95)
res_all <- bind_cols(resp, resi, ress) %>%
  select(-index1, -type1, -index2, -type2) %>%
  mutate(indexF = factor(index, c("holes", "convex", "skinny", "stringy", "dcor2D", "splines2D", "MICe", "TICe"))) %>%
  arrange(indexF) %>%
  select(-indexF, -index) %>%
  add_column(index = c("holes", "", "convex", "", "skinny", "", "stringy", "", "dcor2D", "", "splines2D", "", "MICe", "", "TICe", "")) %>%
  select(index, everything())
# Good resource for doing tables
# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_pdf.pdf
kable(res_all, format = "latex",  caption = "Comparison of index values between noise projections and structured projections for sample size 1000, using 5th and 95th percentiles from 100 simulated sets. Most indices have much larger values for the structured projections, except for convex.", 
      col.names=c("Index","", "lower", "upper", "lower", "upper", "lower", "upper"),
      digits=2, booktabs = T) %>% 
      add_header_above(c(" ", " ", "pipe" = 2, "sine" = 2, "spiral"=2)) %>%
  group_rows("", 1, 2, latex_gap_space = "-0.2cm") %>%
  group_rows("", 3, 4, latex_gap_space = "-0.2cm") %>%
  group_rows("", 5, 6, latex_gap_space = "-0.2cm") %>%
  group_rows("", 7, 8, latex_gap_space = "-0.2cm") %>%
  group_rows("", 9, 10, latex_gap_space = "-0.2cm") %>%
  group_rows("", 11, 12, latex_gap_space = "-0.2cm") %>%
  group_rows("", 13, 14, latex_gap_space = "-0.2cm") %>%
  group_rows("", 15, 16, latex_gap_space = "-0.2cm") %>% 
  #row_spec(1:16, font_size = 10) %>%
  kable_styling(position = "center") #, font_size = 7) 
```

For simplicity, in the investigations of the index behavior, we fix $p=6$, which corresponds to two independent 2D planes containing nuisance distributions, and one 2D plane containing the structured distribution.
The structured projection is in variables 5 and 6 ($x_5, x_6$). Two samples sizes are used: $n=(100, 1000)$. All variables are standardized to have mean 0 and standard deviation 1. Figure \ref{fig:data} shows samples from each of the families, of the nuisance and structured pairs of variables. Table \ref{tab:indexTable} compares the PPIs for structured projections against those for nuisance variables, based on 100 simulated data sets of each type, using sample size 1000. The lower and upper show the 5th and 95th percentile of values. The holes index is sensitive only to the pipe distribution. All other indexes, except convex show distinctly higher values for the structured projections. The convex index shows the inverse scale to other indices, thus (1-convex) will be used in the assessment of performance of PPIs. The scale for the holes index in its original implementation is smaller than the others ranging from about 0.7 through 1, so it is re-scaled in the performance assessment so that all indices can be plotted on a common scale of 0-1 (details in Appendix). Similarly, the TIC index is re-scaled depending on sample size.

@@

For distributions and indexes with smooth behavior and good squintability, method 3 is the most effective method for optimization. If these two criteria are not met the method may still be useful, but only given an informed starting projection. In such cases we can follow a method similar to that proposed by Friedman [@f87]: we break the optimization in two distinct steps. A first step ("scouting") uses method 2 with large search window and no cooling as a way of stepping over fluctuations and local maxima and yielding an approximation of the global maximum. Note that this likely requires large number of tries, especially as dimension increases, since most randomly picked planes will not be interesting. The second step uses method 3 starting from the approximate maximum, which will take small steps to refine the result to be closer to the global maximum.

@@

The final views obtained in each of the two stages are compared in the Appendix, see Fig. \ref{fig:testpipe}.

@@

## Speed
Examining the computing time as a function of sample size we find that scagnostics and splines2d are fast even for large samples, while all other index functions slow rapidly with increasing sample size. Detailed results are shown in Appendix \ref{app:speed}.

## Parameter choices
Some PPIs have a choice of parameters, and the choice can have an effect on function smoothness, and sensitivity to structure. In Appendix \ref{app:param} we examine the dependence of the scagnostics indexes on the binning, showing that even with small number of bins the indexes are noisy, while they lose ability to see structure. Sensitivity to the number of bins in the MIC index is also examined, showing that tuning the parameter can improve the final result.

## Index enhancement
We identified two potential improvements. First, the issue of noisy index functions may be addressed via smoothing, and we explore different smoothing options for the examples of the skinny and stringy index in Appendix \ref{app:enh}. In addition, rotation dependent indexes may be enhanced by redefining them in a rotation invariant way.

@@

Figure \ref{fig:neutronStarSPLOM} in the Appendix shows the scatterplot matrix (with transparency) of the six parameters. (In astrophysics, scatterplot matrices are often called "corner plots" [@corner].) The diagonal shows a univariate density plot of each parameter, and the upper triangle of cells displays the correlation between variables. From this it can be seen that m1 and m2 are strongly, and slightly, nonlinearly associated. Between the other variables we observe some linear association (R1, R2), some nonlinear association (L1, L2, R1, R2),  heteroskedastic variance in most variables and some bimodality (R1, L1, L2, m1, m2).

@@

Figure \ref{fig:bbhSimulation} in the Appendix shows a scatterplot matrix, of nine of the twelve parameters. (Parameter m2 is not shown because it is strongly linearly associated with m1, phi$\_$jl and psi are not shown because they are uniform, and not associated with other parameters.) Among the nine plotted parameters, strong nonlinear relationships can be seen between the parameters ra, dec and time. The first two describe the position of the event in the sky, and time is the merging time (in GPS units). Because of the elliptical relationship between dec and time, the TIC index is used for PP, even though it is slow to compute. Between the other parameters, the main structure seen is multimodality and some skewness. These patterns are representative of the likelihood function, since most priors are flat, or built to capture growth with volume rather than distance.

@@

All seven PP indexes are applied to the data. Figure \ref{fig:bbhGuided} showing the projections that maximize three of the indexes. TIC and splines2d indexes identify very similar projections, that are based on the three parameters, dec, time and ra. This is to be expected based on the pairwise scatterplots. On the other hand, the 1-convex index finds a very different view, but this is because the optimization doesn't adequately reach a maximum for this index.

@@

The variables time, dec and ra are dropped from the data, and PP is applied to the remaining 8D space. Figure \ref{fig:bbhGuided2} shows the projections which maximize the TIC, splines2D and 1-convex indices. The results provide similar information as already learned from the scatterplot matrix (Figure \ref{fig:bbhSimulation}). The parameters chi$\_$tot and chi$\_$p are linearly related (TIC maxima), and theta$\_$jn has a bimodal distribution yielding the figure 8 shape found by the splines2d index. The 1-convex index finds nothing interesting.

>>>>

abstract: |
  Projection pursuit is used to find interesting low-dimensional projections of high-dimensional data by optimizing an index over all possible projections. Most indexes have been developed to detect departure from known distributions, such as normality, or to find separations between known groups. Here, we are interested in finding projections revealing potentially complex bivariate patterns, using new indexes constructed from scagnostics and a maximum information coefficient, with a purpose to detect unusual relationships between model parameters describing physics phenomena. The performance of these indexes is examined with respect to ideal behaviour, using simulated data, and then applied to problems from gravitational wave astronomy. The implementation builds upon the projection pursuit tools available in the R package, tourr, with indexes constructed from code in the R packages, binostics, minerva and mbgraphic.

@@

@@

The driving application is from physics, to aid the interpretation of model fits on experimental results. A physical model can be considered to be a set of $p$ free parameters, that cannot be measured directly and are determined by fitting a set of $q~ (p<q)$ experimental observations, for which predictions can be made once the $p$ parameters are estimated. (Note here, that while we may have analytic expressions for the predictions, this is not always the case and we often have to rely on numerical computation.) Different sets of model parameters ($n$) found to be  compatible with the experimental results within a selected level of confidence yield the data to be examined using projection pursuit. A single prediction can be a complicated function of all of the free parameters, and typically $q \in [100,1000]$ and $p \sim 10$. Current practice is to examine pairs of parameters, or combinations produced by intuition or prior knowledge. This begs the question, whether important nonlinear associations are missed because they are hidden in linear combinations of more than two variables.

@@

To find appropriate projections pursuit indexes for detecting nonlinear dependencies, the literature on variable selection was a starting point. With high-dimensional data, even plotting all pairs of variables can lead to too many plots, which is what "scagnostics" (@scag, @WW08) were developed to address by providing metrics from which to select the most interesting variable pairs. There are eight scagnostics, of which three ("convex", "skinny" and "stringy") are used here. The question is whether these can be adapted into projection pursuit indexes, to search for unusual features in two-dimensional projections of high-dimensional data. Recent PhD research by @Grimm2016 explored the behavior of scagnostics for selecting variables, and proposed two more that have nicer properties, based on smoothing splines and distance correlation. In addition, two more indexes for measuring dependence have been proposed in the machine learning literature, based on information criteria, maximal and total information coefficient (MIC and TIC) [@Reshef1518], with computationally more efficient versions (MIC_e, TIC_e) [@JMLRv1715308]. These are related to original 1D projection pursuit indexes based on entropy (e.g. @huber85, @JS87). This provides seven current indexes for measuring dependence between two variables, and each is available in an R [@rref] package: binostics [@HWscagR], mbgraphic [@mbgraphic] and minerva [@minerva].

@@

This paper is structured as follows. Section \ref{sec:construct} discusses index construction, and how they can be used in the guided tour. Section \ref{sec:investigate} investigates the behavior of the indexes, explored primarily using tour methods. The new guided tour with these indexes is applied to two examples from gravitational wave astronomy (Section \ref{sec:phys}). The latter two parts are connected in that the application of the new indexes to these problems is the main motivation for the paper, and the simulation study, in the first part, was conducted to better understand the behavior of the indexes in general. The techniques in Section \ref{sec:investigate} define procedures that will be generally useful for researchers developing new projection pursuit indexes to visually assess their behavior. Visual methods to diagnose the index behavior is important because PP is primarily used for visualization.  The paper finishes with a discussion about the limitations of this work, and the potential future directions.

@@

A comparison between these indexes for the **purpose of variable selection**, but not projection pursuit, was discussed in @Grimm2016. It is only available in German, we we summarize the main findings here. The scagnostics measures are flexible, and calculating the full set of measures provides useful guidance in variable selection. However, they are found to be highly sensitive to outlying points and sample size (as a consequence of the binning). Both splines2D and dcor2D are found to be robust in this respect, but splines2D is limited to functional dependence, while dcor2D is found to take large values only in scenarios with large linear correlation. The mutual information based index functions (MIC, TIC) are found to be flexible, but are sensitive to the sample size and often take relatively large values even when no association is present. A brief comparison of MIC and dcor2D was also provided in @2014arXiv1401.7645S.

In addition to the seven indexes described above, we will also include the holes index available in the `tourr` package, see (@CBC93, @Cook:2007:IDG:1557227). This serves as a benchmark, demonstrating some desired behavior. The index takes maximum values for a central hole in the distribution of the projected data.

@@

if(!file.exists("cache/indexTable.rda")){
  set.seed(505)
  res <- tableTemp()
  for(ds in c("Pipe", "Sine", "Spiral")){
    for (size in c(100, 1000)){
      for (type in c("structure", "noise")){
        res <- bind_rows(res,getIdx(ds, size, type))
      }
    }
  }
  res <- res %>% arrange(data, size, index, type) %>% select(data, size, index, type, value5, value95)
  save(res, file = "cache/indexTable.rda")
} else {
  load("cache/indexTable.rda")
}
#rescale holes index first
res <- res %>%
  rowwise() %>%
  mutate(value5 = if_else(index=="holes", rescaleHoles(value5), value5)) %>%
  rowwise() %>%
  mutate(value95 = if_else(index=="holes", rescaleHoles(value95), value95))
resp <- res %>% filter(size == 1000, data == "Pipe") %>%
  select(index, type, value5, value95) %>%
  rename(lower=value5, upper=value95)
resi <- res %>% filter(size == 1000, data == "Sine") %>%
  select(index, type, value5, value95) %>%
  rename(lower=value5, upper=value95)
ress <- res %>% filter(size == 1000, data == "Spiral") %>%
  select(index, type, value5, value95) %>%
  rename(lower=value5, upper=value95)
res_all <- bind_cols(resp, resi, ress) %>%
  select(-index1, -type1, -index2, -type2) %>%
  mutate(indexF = factor(index, c("holes", "convex", "skinny", "stringy", "dcor2D", "splines2D", "MICe", "TICe"))) %>%
  arrange(indexF) %>%
  select(-indexF, -index) %>%
  add_column(index = c("holes", "", "convex", "", "skinny", "", "stringy", "", "dcor2D", "", "splines2D", "", "MIC", "", "TIC", "")) %>%
  select(index, everything())
# Good resource for doing tables
# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_pdf.pdf
kable(res_all, format = "latex",  caption = "Comparison of index values between noise projections and structured projections for sample size 1000, using 5th and 95th percentiles from 100 simulated sets. Most indexes have much larger values for the structured projections, except for convex.", 
      col.names=c("Index","", "lower", "upper", "lower", "upper", "lower", "upper"),
      digits=2, booktabs = T) %>% 
      add_header_above(c(" ", " ", "pipe" = 2, "sine" = 2, "spiral"=2)) %>%
  group_rows("", 1, 2, latex_gap_space = "-0.2cm") %>%
  group_rows("", 3, 4, latex_gap_space = "-0.2cm") %>%
  group_rows("", 5, 6, latex_gap_space = "-0.2cm") %>%
  group_rows("", 7, 8, latex_gap_space = "-0.2cm") %>%
  group_rows("", 9, 10, latex_gap_space = "-0.2cm") %>%
  group_rows("", 11, 12, latex_gap_space = "-0.2cm") %>%
  group_rows("", 13, 14, latex_gap_space = "-0.2cm") %>%
  group_rows("", 15, 16, latex_gap_space = "-0.2cm") %>% 
  #row_spec(1:16, font_size = 10) %>%
  kable_styling(position = "center") #, font_size = 7) 
```

For simplicity, in the investigations of the index behavior, we fix $p=6$, which corresponds to two independent 2D planes containing nuisance distributions, and one 2D plane containing the structured distribution.
The structured projection is in variables 5 and 6 ($x_5, x_6$). Two samples sizes are used: $n=(100, 1000)$. All variables are standardized to have mean 0 and standard deviation 1. Figure \ref{fig:data} shows samples from each of the families, of the nuisance and structured pairs of variables. Table \ref{tab:indexTable} compares the PPIs for structured projections against those for nuisance variables, based on 100 simulated data sets of each type, using sample size 1000. The lower and upper show the 5th and 95th percentile of values. The holes index is sensitive only to the pipe distribution. All other indexes, except convex show distinctly higher values for the structured projections. The convex index shows the inverse scale to other indices, thus (1-convex) will be used in the assessment of performance of PPIs. The scale for the holes index in its original implementation is smaller than the others ranging from about 0.7 through 1, so it is re-scaled in the performance assessment so that all indices can be plotted on a common scale of 0-1 (details are given in the Appendix). Similarly, the TIC index is re-scaled depending on sample size.

@@

For distributions and indexes with smooth behavior and good squintability, method 3 is the most effective method for optimization. If these two criteria are not met the method may still be useful, but only given an informed starting projection. In such cases we can follow a method similar to that proposed by @f87: we break the optimization in two distinct steps. A first step ("scouting") uses method 2 with large search window and no cooling as a way of stepping over fluctuations and local maxima and yielding an approximation of the global maximum. Note that this likely requires large number of tries, especially as dimension increases, since most randomly picked planes will not be interesting. The second step uses method 3 starting from the approximate maximum, which will take small steps to refine the result to be closer to the global maximum.

@@

The final views obtained in each of the two stages are compared in the Appendix.

@@

## Speed
Examining the computing time as a function of sample size we find that scagnostics and splines2d are fast even for large samples, while all other index functions slow rapidly with increasing sample size. Detailed results are shown in the Appendix.

## Parameter choices
Some PPIs have a choice of parameters, and the choice can have an effect on function smoothness, and sensitivity to structure. In the Appendix we examine the dependence of the scagnostics indexes on the binning, showing that even with small number of bins the indexes are noisy, while they lose ability to see structure. Sensitivity to the number of bins in the MIC index is also examined, showing that tuning the parameter can improve the final result.

## Index enhancement
We identified two potential improvements. First, the issue of noisy index functions may be addressed via smoothing, and we explore different smoothing options for the examples of the skinny and stringy index in the Appendix. In addition, rotation dependent indexes may be enhanced by redefining them in a rotation invariant way.

@@

A scatterplot matrix (with transparency) of the six parameters is shown in the Appendix. (In astrophysics, scatterplot matrices are often called "corner plots" [@corner].) The diagonal shows a univariate density plot of each parameter, and the upper triangle of cells displays the correlation between variables. From this it can be seen that m1 and m2 are strongly, and slightly, nonlinearly associated. Between the other variables we observe some linear association (R1, R2), some nonlinear association (L1, L2, R1, R2),  heteroskedastic variance in most variables and some bimodality (R1, L1, L2, m1, m2).

@@

A scatterplot matrix, of nine of the twelve parameters, is shown in the Appendix. (Parameter m2 is not shown because it is strongly linearly associated with m1, phi$\_$jl and psi are not shown because they are uniform, and not associated with other parameters.) Among the nine plotted parameters, strong nonlinear relationships can be seen between the parameters ra, dec and time. The first two describe the position of the event in the sky, and time is the merging time (in GPS units). Because of the elliptical relationship between dec and time, the TIC index is used for PP, even though it is slow to compute. Between the other parameters, the main structure seen is multimodality and some skewness. These patterns are representative of the likelihood function, since most priors are flat, or built to capture growth with volume rather than distance.

@@

All seven PP indexes are applied to the data. Figure \ref{fig:bbhGuided} shows the projections that maximize three of the indexes. TIC and splines2d indexes identify very similar projections, that are based on the three parameters, dec, time and ra. This is to be expected based on the pairwise scatterplots. On the other hand, the 1-convex index finds a very different view, but this is because the optimization doesn't adequately reach a maximum for this index.

@@

The variables time, dec and ra are dropped from the data, and PP is applied to the remaining 8D space. Figure \ref{fig:bbhGuided2} shows the projections which maximize the TIC, splines2D and 1-convex indices. The results provide similar information as already learned from the scatterplot matrix. The parameters chi$\_$tot and chi$\_$p are linearly related (TIC maxima), and theta$\_$jn has a bimodal distribution yielding the figure 8 shape found by the splines2d index. The 1-convex index finds nothing interesting.
